Comparative analysis of PMV Models accuracy implemented in the ISO 7730:2005 and ASHRAE 55:2023

### 2025-01-17 Manuscript/src/sections/conclusions.tex
review: updated the whole paper with the new results I_cl_r ISO.

The overall accuracy and F1-macro score of the \ac{pmv} is \qty{33}{\percent} and \num{.16}, respectively.
While the overall accuracy and F1-macro score of the \ac{pmv-ce} is \qty{34}{\percent} and \num{.16}, respectively.
Both models have an accuracy lower or equal to \qty{7}{\percent} when predicting either `warm,' `hot,' or `cold.'
This value is lower than random guessing (\qty{14}{\percent}).
This means that the models are misleading the users.
When only comparing the results of the two models for entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
For example, the median bias of the \ac{pmv} model is \var{bias_median_pmv_0.2} while the median bias for the \ac{pmv-ce} is \var{bias_median_pmv_ce_0.2}.

@@

The very low prediction accuracy of both model formulations, along with the lack of data at the extremes of their applicability limits, suggests that these limits should be reconsidered.
Based on our results, we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
In other words, we only recommend using it for determining whether an environment may be deemed thermally neutral by a large group of occupants.
The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classify if individuals are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988} provide corroborating evidence that supports our results.

We also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} standards.
Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
\todo{But this statement is incorrect for ASHRAE Std 55, which only defines its single comfort zone using +/-0.5}
Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
\todo{unclear what limits you are talking about, or who the audience is for your suggestions.  The ComfortTool allows extrapolation to high PMVs, but that is not true for ASHRAE 55 itself.  Your suggestions are actually just aimed at yourselves as keepers of the ComfortTool}
The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
Consequently, we suggest reducing both standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
\begin{table}[htb!]
    \centering
    \begin{tabular}{cc}
        \toprule
        Variable & New Proposed range \\
        \midrule
        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
        \ac{rh} & \qtyrange{10}{80}{\percent} \\
        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
        \ac{met} & \qtyrange{1}{2}{met} \\
        \ac{pmv} & \numrange{-.5}{.5}{} \\
        \bottomrule
    \end{tabular}
    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} standards.}
    \label{tab:ranges}
\end{table}
\mycite{Humphreys2002} reached a similar conclusion and observed that the ranges of inputs consistent with a valid use of PMV are much narrower than those given in \gls{7730}.
The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
We then recommend that the \gls{55} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.
\todo{and why not the ISO Standard also?  After all, ASHRAE was copying ISO on that one}

@@

>>>>

% todo check values manually
The overall accuracy and F1-macro score of the \ac{pmv} is \var{overall_acc_pmv} and \num{.16}, respectively.
While the overall accuracy and F1-macro score of the \ac{pmv-ce} is \var{overall_acc_pmv_ce} and \num{.15}, respectively.
% todo check values manually
Both models have an accuracy lower or equal to \qty{5}{\percent} when predicting either `warm', `hot', or `cold'.
This value is lower than random guessing (\qty{14}{\percent}).
This means that the models are misleading the users.
% todo check values manually
When only comparing the results of the two models for entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} the \ac{pmv} outperformed the \ac{pmv-ce} on different metrics.
For example, the median bias of the \ac{pmv} model is \var{bias_median_pmv_0.2} while the median bias for the \ac{pmv-ce} is \var{bias_median_pmv_ce_0.2}.

@@

The very low prediction accuracy of both model formulations, along with the lack of data at the extremes of their applicability limits, suggests that these limits should be reconsidered.
Based on our results, we recommend restricting the use of the \ac{pmv} only to values $\lvert \textrm{PMV}\lvert \leq 0.5$.
In other words, we only recommend using it for determining whether an environment may be deemed thermally neutral by a large group of occupants.
The model, however, is not capable of determining the degree by which people are dissatisfied and can only be used to classify if individuals are on the `cold' (\ac{pmv}~$\leq 0.5$) or `warm' (\ac{pmv}~$\geq 0.5$) side.
Several peer-reviewed papers \cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988} provide corroborating evidence that supports our results.
We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption.
However, this is not the case.
As shown in Figure~\ref{fig:comfort_regios_pmv_pmvce} increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office settiong to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
Hence, we would like to point out that it is not the \ac{pmv} model that is a limiting factor in saving energy, but instead is that in buildings we cool the air rather than moving it.
A field study in Singapore showed that allowing participants to control \ac{v} while simultaneously increasing \ac{tdb} from \qty{24}{\celsius} to \qty{26.5}{\celsius} reduced the energy consumption by \qty{32}{\percent}~\cite{kent_energy_2023}.

We also recommend restricting the ranges on both environmental and personal factors set by both \gls{55} and \gls{7730} standards.
Currently, these limits extend far beyond the range of data contained in the \ac{db2} dataset.
\todo{But this statement is incorrect for ASHRAE Std 55, which only defines its single comfort zone using +/-0.5}
Being the \ac{db2} the largest thermal comfort dataset in the world and containing data from all continents, we deem it not to be necessary to have such wide applicability limits for both \gls{55} and \gls{7730} standards when these environmental variables are rarely observed in buildings across the world.
\todo{unclear what limits you are talking about, or who the audience is for your suggestions.  The ComfortTool allows extrapolation to high PMVs, but that is not true for ASHRAE 55 itself.  Your suggestions are actually just aimed at yourselves as keepers of the ComfortTool}
The lack of data available beyond the ranges depicted in Figure~\ref{fig:dist_input_data} did not allow us to test the accuracy of both the \ac{pmv} and \ac{pmv-ce} above those ranges.
Consequently, we suggest reducing both standards' applicability limits to the ranges shown in Table~\ref{tab:ranges}, until more data are collected.
\begin{table}[htb!]
    \centering
    \begin{tabular}{cc}
        \toprule
        Variable & New Proposed range \\
        \midrule
        \ac{tdb} & \qtyrange{17}{30}{\celsius} \\
        \ac{tr} & \qtyrange{17}{30}{\celsius} \\
        \ac{rh} & \qtyrange{10}{80}{\percent} \\
        \ac{clo} & \qtyrange{0.3}{1.5}{clo} \\
        \ac{met} & \qtyrange{1}{2}{met} \\
        \ac{pmv} & \numrange{-.5}{.5}{} \\
        \bottomrule
    \end{tabular}
    \caption{New proposed applicability limits for the \gls{55} and \gls{7730} standards.}
    \label{tab:ranges}
\end{table}
\mycite{Humphreys2002} reached a similar conclusion and observed that the ranges of inputs consistent with a valid use of PMV are much narrower than those given in \gls{7730}.
The limit on \ac{met} is also corroborated by evidence provided by \mycite{doherty_evaluation_1988}.
We then recommend that both the \gls{55} and \gls{7730} should change back the upper value of \ac{met} to \qty{2}{met} as it was in the 2020 version of the standard.

@@

Finally, we would like to highlight that the current definition of the \ac{pmv} model ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
Both the \gls{55} and \gls{7730} standards should clarify the minimum number of occupants required to use the \ac{pmv} model.
Can it be used for a space with only one occupant?

### 2025-01-17 Manuscript/src/sections/introduction.tex
review: updated the whole paper with the new results I_cl_r ISO.

            \draw [arrow] (start) -- (dec1);
            \draw [arrow] (dec1) -- node[above, pos=0.3] {Yes} (pro1a);
            \draw [arrow] (dec1) -- node[right, pos=0.3] {No} (dec2);
            \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
            \draw [arrow] (dec2) -- (dec3);
        \end{tikzpicture}
        \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{7730} standard.
            $\alpha$ is the correction coefficient for \acs{clo}.
            Since the \gls{7730} does contain some errors and it does not provide all the necessary information to calculate $\alpha$ we used the equations provided in the \gls{9920} to derive the value of $\alpha$.
        In accordance with the \gls{7730} and \gls{9920} standards, since we did not have information about the walking speed of the participants, we estimated it using the provided equation.
        }
        \label{fig:flowchart_pmv_iso}
    \end{subfigure}
    \caption{Flowcharts depicting the steps for the calculation of the \ac{pmv} following the \gls{55} and \gls{7730} standards.}
    \label{fig:flowchart_pmv_calculation}
\end{figure}
In summary, as shown in Figure~\ref{fig:flowchart_pmv_ce} when the \ac{vr} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
\ac{vr} is different from the \ac{v}.
According to the standard, the measured value should be adjusted using this equation \ac{vr} = \ac{v}$+ 0.3 ($\acs{met}$-1)$ if \ac{met}~$>$~\qty{1}{met}.
The standard also prescribes the use of \ac{clor} and not of \ac{clo} as an input.
For \ac{met} $> \qty{1.2}{met}$, \acs{clor} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$.
The value of \ac{ce} is calculated using the \ac{set} equation.
The \ac{ce} is then subtracted from both the \ac{tdb} and \ac{tr}.
The resulting values become the new inputs in the \ac{pmv} model.
Since the \ac{ce} accounts for convective and evaporative heat losses from the person to its environment, the value of \ac{vr}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
The other three input parameters (\ac{clor}, \ac{rh}, \ac{met}) remain unchanged.
% todo this is no longer true and it should be removed or rephrased
Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ only when the value of \ac{vr} is higher than \qty{0.1}{\m\per\s}.
To illustrate the extent of the differences between the outputs of the two models, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{vr} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clor}~=~\qty{0.5}{clo}.
We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
\todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. There are plenty of lab studies that have measured results in that region.}
\todo[color=blue!40]{FT. I can include a Figure for 0.6m/s but I did not since based on the ASHRAE DB 2 we hardly reach that value in buildings.}
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models for two values of \ac{vr}.
    We set \ac{met}~=~\qty{1.2}{met}, \ac{clor}~=~\qty{0.5}{clo}.
    \label{fig:comfort_regios_pmv_pmvce}}
\end{figure}
The results show that for \ac{vr}~=~\qty{0.4}{\m\per\s} and \ac{rh}~=~\qty{50}{\percent} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} and shifted towards warmer temperatures.
For \ac{vr}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.

The main rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds ~\cite{huang_applicability_2014}.
The \ac{pmv} does not accurately estimate evaporative heat losses trough sweating because it assumes that sweating is constant and only varies as a function of \ac{met}.
Hence, if sweating is required to maintain the body in thermal equilibrium, the \ac{pmv} model assumes that the body would gain heat.
The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
\todo[inline]{true enough, but in the standard the objective is to have a rational physiology model.
The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
\todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}

Our analysis has, however, revealed several limitations specific to the \ac{pmv-ce} model, as outlined below.

% \mycite{huang_applicability_2014}, using 19 averaged data points obtained from a set of experiments involving 30 subjects, showed the \ac{set} model could be used to estimate the thermal sensation with two different ways of estimating the airspeed. 
% They developed a model to relate the SET and the TSV. 
% They also plotted the \ac{pmv} versus the TSV and provided a visual analysis of its performance. They concluded that the SET is better than the \ac{pmv} model in predicting thermal sensation votes when subjects are exposed to air movements, but they did not report the performance of the \ac{pmv} model.
% Moreover, they did not compare the accuracy of the \ac{pmv-ce} with the \ac{pmv} model.

@@

    \item Although the \ac{ce} claims to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model.

@@

    However, these changes were not based on any scientific evidence and significantly affect the output of the model.

@@

    \todo[inline]{This may or may not be significant. If it is significant, the process could be simplified to a total of three calculations while remaining a rational method. What’s in there now was Tyler’s (CBE’s) choice. Charlie and I have used it in the simpler way. We could propose this to the Standards Committee if we wanted to. We would have to calculate two validation tables for the two approaches.  }

>>>>

            \draw [arrow] (start) -- (dec1);
            \draw [arrow] (dec1) -- node[above, pos=0.3] {Yes} (pro1a);
            \draw [arrow] (dec1) -- node[right, pos=0.3] {No} (dec2);
            \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
            \draw [arrow] (dec2) -- (dec3);
        \end{tikzpicture}
        \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{7730} standard.
            $\alpha$ is the correction coefficient for \acs{clo}.
            Since the \gls{7730} does contain some errors and it does not provide all the necessary information to calculate $\alpha$ we used the equations provided in the \gls{9920} to derive the value of $\alpha$.
        The \gls{7730} standard, also provides an equation to calculat the walking speed if this information is not available.
        }
        \label{fig:flowchart_pmv_iso}
    \end{subfigure}
    \caption{Flowcharts depicting the steps for the calculation of the \ac{pmv} following the \gls{55} and \gls{7730} standards.}
    \label{fig:flowchart_pmv_calculation}
\end{figure}
In summary, as shown in Figure~\ref{fig:flowchart_pmv_ce}, when the \ac{vr} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
\ac{vr} is different from the \ac{v}.
According to the standard, the measured value should be adjusted using this equation \ac{vr} = \ac{v}$+ 0.3 ($\acs{met}$-1)$ if \ac{met}~$>$~\qty{1}{met}.
The standard also prescribes the use of the \ac{clor} as an input rather than the \ac{clo}.
For \ac{met} $> \qty{1.2}{met}$, \acs{clor} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$.
The value of \ac{ce} is calculated using the \ac{set} equation.
The \ac{ce} is then subtracted from both the \ac{tdb} and \ac{tr}.
The resulting values become the new inputs in the \ac{pmv} model.
Since the \ac{ce} accounts for convective and evaporative heat losses from the person to its environment, the value of \ac{vr}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
The other three input parameters (\ac{clor}, \ac{rh}, \ac{met}) remain unchanged.

Figure~\ref{fig:flowchart_pmv_iso} shows the calculation routine for the \ac{pmv} model as per the \gls{7730} standard.
The same equation as in the \gls{55} is used to adjust the \ac{vr} value.
However, the \ac{clor} is calculated differently.
In the \gls{7730}, the \ac{clor} varies as a function of \ac{clo}, \ac{met}, walking speed, and \ac{vr}.

Consequently, while the two \ac{pmv} formulations mainly differ when the value of \ac{vr} is higher than \qty{0.1}{\m\per\s}.
It should be noted that the two models also use different equations to calculate the \ac{clor}.

To illustrate the extent of the differences between the outputs of the two models, only as a function of \ac{vr}, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) assuming the same \ac{met}~=~\qty{1.2}{met} and \ac{clor}~=~\qty{0.5}{clo}.
We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
\todo{It would be great to add 0.6m/s to this figure. That would be the objective for ceiling fan designs. There are plenty of lab studies that have measured results in that region.}
\todo[color=blue!40]{FT. I can include a Figure for 0.6m/s but I did not since based on the ASHRAE DB 2 we hardly reach that value in buildings.}
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models for two values of \ac{vr}.
    We set \ac{met}~=~\qty{1.2}{met}, \ac{clor}~=~\qty{0.5}{clo}.
    \label{fig:comfort_regios_pmv_pmvce}}
\end{figure}
The results show that for \ac{vr}~=~\qty{0.4}{\m\per\s} and \ac{rh}~=~\qty{50}{\percent} the comfort regions estimated using the \ac{pmv-ce} is \qty{0.7}{\celsius} wider than the one calculated using the \ac{pmv} and shifted towards warmer temperatures.
For \ac{vr}~=~\qty{0.2}{\m\per\s}, the same difference is \qty{0.4}{\celsius}.

The main rationale for the development of the \ac{pmv-ce} model is that the original \ac{pmv} formulation does not accurately estimate convective and evaporative heat losses from the skin to the environment under elevated air speeds~\cite{huang_applicability_2014}.
The \ac{pmv} does not accurately estimate evaporative heat losses trough sweating because it assumes that sweating is constant and only varies as a function of \ac{met}.
Hence, if sweating is required to maintain the body in thermal equilibrium, the \ac{pmv} model assumes that the body would get hotter.
The human body, on the other hand, uses sweating to maintain a constant body temperature in `warm' environments.
\todo[inline]{true enough, but in the standard the objective is to have a rational physiology model.
The Gagge models are as well or better documented than the Fanger heat balance model, and have been available to Standards users in the ASHRAE Handbook of Fundamentals.}
\todo[inline, color=blue!40]{FT. The PMV CE and the Gagge model are very different, hence, publications about the latter do not justify the implementation proposed in the Std 55}

Despite this claim from the authors of the \ac{pmv-ce}, our analysis has, however, revealed several limitations specific to the \ac{pmv-ce} model, as outlined below.

@@

    \item Although the \ac{ce} claims to better account for convective heat losses from the skin to the environment due to the air movement, the \ac{pmv-ce} model uses the same underlying equations as the \ac{pmv} model, hence it has the same limitations.

@@

    However, these changes were not based on any scientific evidence, despite significantly affecting the output of the model.

@@

    \todo[inline]{This may or may not be significant. If it is significant, the process could be simplified to a total of three calculations while remaining a rational method. What’s in there now was Tyler’s (CBE’s) choice. Charlie and I have used it in the simpler way. We could propose this to the Standards Committee if we wanted to. We would have to calculate two validation tables for the two approaches.}

### 2025-01-17 Manuscript/src/sections/methodology.tex
review: updated the whole paper with the new results I_cl_r ISO.

\section{Methodology}\label{sec:methodology}
The \ac{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world, and it is the largest of its kind~\cite{FoldvaryLicina2018, db2dryad}.
\Ac{tsv} were collected using a right-here-right-now survey and environmental parameters were measured and logged in the proximity of participants when they completed the survey.
The \ac{db2} only contains data collected by researchers in cross-sectional field studies that have been published in peer-reviewed journals.

\subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
This filtered out approximately \var{entries_db_valid} of the data.
In our analysis, we used the value \ac{vr} which is adjusted as a function of \ac{met} and not the air speed measured by the researchers.
This is because both \ac{pmv} formulations use the \ac{vr} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.

@@

Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
However, since the thermal sensation was measured with a seven-point scale and the \ac{pmv} has no upper or lower boundary, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.

In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same methodology.
For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
This could introduce a bias in the measurements and, consequently, in the results of the models.
Moreover, since the two \ac{pmv} formulations only differ when \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
\begin{enumerate}[ {}1{)} ]
    \item all the data points with \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv};
    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s};
    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured air speed at three different heights.
    These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
\end{enumerate}
Comparing the results from these three datasets will allow us to assess the impact of the data quality on the results of the models.
By only including those studies that measured air speed at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards.
We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was changed to \qty{0.1}{\m\per\s} only in the \gls{55}.
The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
The threshold in the ASHRAE standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.

@@

\subsection{Model Validation}\label{subsec:model-validation}
The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants, but it is commonly used both in singly occupied zones and in areas with several hundred people.
We first grouped participants' responses by their \ac{tsv}, and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
This approach is useful for performing an overall assessment of the accuracy of the models, but it introduces an error due to the rounding of both the \ac{pmv}, \ac{pmv-ce}, and \ac{tsv} values.
The latter were sometimes collected using a continuous scale and not a discrete one, in other words, they could take any value in the range from -3.0 to 3.0.

To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002}.
These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{vr}).
If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{-0.25} and \num{0.25}.
However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers.
Consequently, we expanded the range to \num{-.5} and \num{0.5}.
This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range and ideally, we would want to have the \ac{pmv} to have higher precision, however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale.
When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.
Grouping the participants' responses by their \ac{tsv} mixes people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.

>>>>

\section{Methodology}\label{sec:methodology}
The \ac{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world, and it is the largest of its kind~\cite{FoldvaryLicina2018, db2dryad}.
\Ac{tsv} were collected using a right-here-right-now survey and environmental parameters were measured and logged in the proximity of participants when they completed the survey.
The \ac{db2} only contains data collected indoors by researchers in cross-sectional field studies that have been published in peer-reviewed journals.

\subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
This filtered out approximately \var{entries_db_valid} of the data.

@@

Fanger and the \gls{7730} state that the \ac{pmv} should only be used when its absolute value is lower than 2~\cite{Fanger1970, iso7730}.
However, since the thermal sensation was measured with a seven-point scale, the \ac{pmv} has no upper or lower boundary, and the \gls{55} does not specify the range of applicability of the \ac{pmv-ce} model, we decided to keep the data that felt within the following ranges $|$\ac{tsv}$|$~$\leq$~\num{3} or $|$\ac{pmv}$|$~$\leq$~\num{3.5}.

In addition, it should also be noted that while all the data included in the \ac{db2} are published in peer-reviewed papers, not all entries were collected using the same standardised methodology.
For example, while both thermal comfort standards recommend measuring \ac{v} at three different heights, this was not always done.
The absence of standardized measurement protocols, the use of inaccurate or un-calibrated sensors, and the use of different reference tables to estimate \ac{clo} and \ac{met} may all introduce an error in the measurements and, consequently, in the results of both models.
This issue has been already evidenced in the literature by \mycite{dambrosio_alfano_role_2011}.
They showed that the reliability of the \ac{pmv} model is strongly affected by the accuracy of the input data.
In our analysis, we compensate this issue by:
\begin{enumerate}[ {}1{)} ]
    \item comparing the results of the two models using the same dataset; hence, the different in the output is mainly due to the different formulations of the models;
    \item using a large dataset to reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards since it can be assumed that measurement errors random across different studies.
    \item using only datasets that have been published in peer-reviewed journals.
%    \item excluding all those studies that did not measure all the environmental parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
\end{enumerate}

In addition, since the two \ac{pmv} formulations mainly differ when \ac{vr}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
\begin{enumerate}[ {}1{)} ]
    \item all the data points with \ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv};
    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s};
    \item the data points in 1) with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
    These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
\end{enumerate}
Comparing the results from these three datasets will allow us to assess the impact of the data quality on the results of the models.
By only including those studies that measured air speed at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by both standards.
We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s} and it was changed to \qty{0.1}{\m\per\s} only in the \gls{55}.
The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume~\cite{zukowska_impact_2012}.
In \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}, we also present the results generated using those entries with \ac{vr}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
We have not included these results here, as they are comparable to those presented in this section and lead to the same conclusions.

@@

\subsection{Model Validation}\label{subsec:model-validation}
The \ac{pmv} model was developed to predict the average thermal sensation of a large group of occupants, but it is often applied in both singly occupied spaces and areas accommodating several hundred people.
Partially because neither of the two standards specifies the minimum number of people needed to apply the model.
We first grouped participants' responses by their \ac{tsv}, and we determined how many of these votes were correctly labelled by the \ac{pmv} and \ac{pmv-ce} models.
This approach is useful for performing an overall assessment of the accuracy of the models, but it introduces an error due to the rounding of both the \ac{pmv}, \ac{pmv-ce}, and \ac{tsv} values.
The latter were sometimes collected using a continuous scale and not a discrete one, in other words, they could take any value in the range from -3.0 to 3.0.

To remove this rounding error, we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
However, on their own, are a low-precision estimate of the overall accuracy of the model~\cite{Humphreys2002} since the model is not expected to predict the exact \ac{tsv} of each participant.
These values were then binned using several independent variables (e.g., \ac{tdb}, \ac{vr}).
If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
The standard deviation would reflect the combined effect of the people's individual differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)~\cite{Humphreys2002}.
This is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.
According to \mycite{Humphreys2002} the model can be considered accurate if the above-mentioned differences are between \num{\pm 0.25}.
However, we believe that this range is too narrow since \ac{tsv} is commonly measured using an ordinal scale of integers and even when environmental parameters are measured within the accuracy of the standards, the output of the model may vary by more than \num{\pm 0.25}~\cite{dambrosio_alfano_role_2011}.
Consequently, we expanded the range to \num{\pm 0.5}.
This is needed to ensure that each \ac{pmv} value can be matched to a \ac{tsv}.
We are aware that an error of \num{.5} is approximately \qty{7}{\percent} of the total range.
Ideally, we would want to have the \ac{pmv} to have higher precision.
however, this is a necessary limitation introduced by the low resolution of the \ac{tsv} scale and by the accuracy of the inputs.
When the aim is to predict the average thermal sensation of a large group of occupants, binning data by the independent variables is a better representation.
Grouping the participants' responses by their \ac{tsv} mixes people who reported the same \ac{tsv} vote but could have been in different locations and moments of the year and day.
